\chapter*{Acknowledgement}
I would like to express my deepest gratitude to my thesis supervisor Dr. Vaibhav Arhode for his invaluable guidance, support, and mentorship throughout this project. His expertise, insightful feedback, and constant encouragement have been instrumental in shaping my understanding and enhancing the quality of my work.
I would also like to extend my heartfelt appreciation to co-supervisor Dr. Kaushal Nishad for his valuable contributions and collaboration. His dedication, knowledge, and enthusiasm have greatly enriched the research process and outcomes.
I am immensely grateful to Anil Ji, our lab assistant, for his tireless efforts in facilitating the smooth functioning of the laboratory. His technical assistance and willingness to go above and beyond have been truly commendable.
I would also like to acknowledge my lab partners for their camaraderie, cooperation, and teamwork. Their support and collective efforts have made this journey both enjoyable and productive.
Lastly, I would like to express my gratitude to my friends who have provided unwavering support, encouragement, and motivation throughout this endeavor.
I am truly fortunate to have had the opportunity to work with such exceptional individuals, and their contributions have been invaluable in the successful completion of this project.

% \lipsum[66]

\vspace*{1.5cm}
\begin{flushright}
\authorname
\end{flushright}


% \vspace*{-1.3cm}
% \noindent
% Place: Indian Institute of Technology Kanpur\\
% Date: \today 
